%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\raceentry{Drow}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\tagline{"Time to die for the Spider Queen."}

The Drow are perhaps the most overused bunch of villains ever. Their entire ability set is one that is supposed to neutralize the advantages of player characters so that characters can have mirror matches against NPC parties without doubling their treasure. With magic items that \textit{turn off} once they are brought out of Drow controlled regions, spell-resistance, and spell-like abilities designed to specifically negate common player-character tactical advantages, they can easily compete with Player Characters with massively more permanent magical equipment. And that means that they can be fought and killed several times without supercharging party treasure.

But if you want to \textit{play} a Drow character, you don't want any of that crap. In fact, if you want a Drow character, probably the maxim you are looking for is "WWDD?" and the answer is probably "Fight with two scimitars." But more than that, there are a number of abilities that Drow characters in stories exhibit that people want. And then there are the game mechanical abilities in the rulebook that the characters in stories obviously don't have (like \textit{Touch of Fatigue}, what's up with that?) So here it is, the LA +0 Drow that people actually want to play:

\begin{itemize*}
\item Medium Size
\item 30' movement.
\item Humanoid Type (Elf subtype)
\item Darkvision 120'
\item +2 Dexterity, -2 Constitution
\item Daylight Sensitivity: While in brightly lit surroundings (such as a \textit{daylight} spell), a Drow suffers a -2 penalty to attack rolls and precision-based skill checks.
\item Drow with a Charisma of at least 10 may cast \textit{deeper darkness} (duration 4 hours), and \textit{fairie fire} as spell-like abilities with a caster level equal to their character level once per day each.
\item +2 bonus to saving throws against spells and spell-like abilities.
\item +2 bonus to Spot, and Listen checks.
\item Drow never sleep and are immune to \textit{sleep} effects. Drow must still 
perform their 4 hour daily trance to stay coherent and rested.
\item Drow live an exceedingly interesting life and every Drow has proficiency 
with the rapier and an exotic ranged weapon of their choice.
\item Favored Classes: Cleric and Wizard
\item Automatic Languages: Elvish
\item Bonus Languages: Abyssal, Beholder, Common, Draconic, Drow Sign Language, Dwarvish, Gnome, Kuo-Toa, Terran, Undercommon. 
\end{itemize*}
