%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\classentry{Dragonfire Adept}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\tagline{"I studied the ways of the dragons. Mostly I learned how to set punks like you on fire."}

Dragonfire Adepts are heroes (and villains) who wander around the world doing draconic things. Mostly this entails breathing on things to set them on fire or whatever. They have studied the powers of dragons - those fierce beasts that \sout{fly around terrorising the countryside} laze about like cats and collect shinies like magpies, the magical creatures that \sout{tower above cities} could actually sleep on your bed, generally. 

Dragonfire Adepts are not born, they are taught. You don't just appear as a dragonfire adept because one of your ancestors needs to go to prison, they actually sit down and decide to learn how to be awesome, like a Wizard or Fighter or Archivist. That said, many of them learn their ways on the job, and a large portion of the learning is always going to be via experience. They are almost compelled to "go out adventuring". 

\textbf{Playing a Dragonfire Adept:} Constitution is awesome for your Hit Points, and more importantly for determining the Save DC of your Breath Weapon. However you can also find uses for Charisma - Spell-Like Abilities are keyed to it, and sometimes you want to talk at people. The other Ability Scores depend very much on what you want to do in particular. Dragonfire Adepts tend to take feats that either make their Breath Weapon more awesome, or that generally make them better at fighting such as Juggernaut and Elusive Target. 

\textbf{Starting Age:} as Cleric 

\textbf{Starting Wealth:} as Rogue 

\textbf{Alignment:} Dragonfire Adepts can be of any alignment, and this often determines the kind of dragon they like best, because they're colour-coded for your convenience. 

\textbf{Hit Die:} d8 

\textbf{Skill Points per level:} 4 + Int 

\textbf{Class Skills:} \linkskill{Appraise}, \linkskill{Bluff}, \linkskill{Climb}, \linkskill{Concentration}, \linkskill{Craft}, \linkskill{Decipher Script}, \linkskill{Diplomacy}, \linkskill{Disable Device}, \linkskill{Gather Information}, \linkskill{Intimidate}, \linkskill{Jump}, \linkskill{Knowledge} (Arcana, Nature, Dungeoneering), \linkskill{Listen}, \linkskill{Ride}, \linkskill{Search}, \linkskill{Sense Motive}, \linkskill{Spellcraft}, \linkskill{Spot}, and \linkskill{Use Magic Device}.

\modebab{}
\goodfor{}
\goodref{}
\goodwil{}

\begin{extraclasstable}{Dragonfire Adept}{\multicolumn{1}{p{1.3cm}}{\textbf{Breath Weapon}}}
\levelone{Blazing Breath, Draconic Mind & 2d6}
\leveltwo{Frosty Breath, Spell-like Abilities I & 3d6}
\levelthree{Projected Breath, Legacy Secret, Dragon Age & 4d6}
\levelfour{Explosive Breath, Arcane Insight & 5d6}
\levelfive{Shocking Breath, Dragoncraft & 6d6}
\levelsix{Sphere Access & 7d6}
\levelseven{Radiating Breath, Scales I & 8d6}
\leveleight{Corrosive Breath & 10d6}
\levelnine{Legacy Secret & 11d6}
\levelten{Booming Breath, Exhalation Enhancement & 12d6}
\leveleleven{Spell-like Abilities II & 13d6}
\leveltwelve{Sphere Access & 15d6}
\levelthirteen{Concussive Breath & 16d6}
\levelfourteen{Scales II, Shadow Breath & 18d6}
\levelfifteen{Legacy Secret & 20d6}
\levelsixteen{Spell-like Abilities III & 23d6}
\levelseventeen{Gaseous Breath & 26d6}
\leveleighteen{Sphere Access, Final Dragon Age & 30d6}
\levelnineteen{Draconic Immunities & 35d6}
\leveltwenty{Annihilating Breath, Sorcerous Secrets & 40d6}
\end{extraclasstable}

\ClassFeatures
All of the following are Class Features of the Dragonfire Adept class:

\textbf{Weapon and Armour Proficiencies:} the Dragonfire Adept is proficient with Light and Medium Armour, but not with Shields of any kind. She is proficient with Simple and Natural Weapons, and one Martial or Exotic Weapon of choice. She is also proficient with any Armour or Shield that is specifically made out of dragon parts, and with any Intelligent Weapon. The Dragonfire Adept can sleep comfortably in any armour made from dragon scales, and treats the Maximum Dexterity as 2 points higher, ignoring the Armour Check Penalty outright. 

\textbf{Blazing Breath (Su):} the Dragonfire Adept has a Breath Weapon. This extends in a Cone out to Close Range, and deals Fire damage with a Reflex Save (Constitution-based) for half. The damage is listed in the table, and is based on the class level, although stuff that advances caster levels or "other class features" should apply to this. This can be used once per round, with a Standard Action, however it can actually be modified by Metabreath Feats, simply adding a delay where there wouldn't normally be one. 

\textbf{Draconic Mind (Ex):} the Dragonfire Adept sees the world in a similar way to how dragons see it. As a result, she gains the [Dragonblooded] Subtype or whatever it is, gains an Insight Bonus to Appraise, Knowledge and Spellcraft checks of +1 per 3 hit dice (round up), and gains Low-Light Vision, and 120' Darkvision. At 6 hit dice, this will become 250' Darkvision and 120' Blindsense and Tremorsense. At 12 hit dice, this becomes 500' Darkvision, 250' Blindsense and Tremorsense, and 120' Blindsight. 

\textbf{Frosty Breath (Su):} at second level, the Dragonfire Adept can elect to deal Cold damage with her breath weapon instead of dealing Fire damage. She can make this choice every time she uses the breath weapon. When using the Cold breath, the save for half is a Fortitude Save. 

\textbf{Spell-Like Abilities I (Sp):} at level two, the Dragonfire Adept can cast the following once each per day: Darkness, Light, Water Breathing, Ventriloquism, Ghost Sound, Command, Fog Cloud, Speak With Animals, Bless, Spider Climb. The caster level equals her hit dice, and the Save DC is 10 + half her hit dice + her Charisma Bonus. 

\textbf{Projected Breath (Su):} at level three, the Dragonfire Adept gains the option of releasing her breath weapon in a Medium Line instead of a Close Cone. 

\textbf{Legacy Secret:} at levels three, nine and fifteen, the Dragonfire Adept may select a Legacy Secret, a special power that is unlocked. Each one provides some special abilities at level three, level nine, and level fifteen.

\begin{itemize*}
\item Dragon Body: the Dragonfire Adept gains a Primary Bite (1d6+Str for a Medium character), a Primary Tail (1d8+Str*1.5 for a Medium character) and two Secondary Claws (1d4+Str*0.5 for a Medium character). These are all actually magic weapons with an Enhancement Bonus of +1 per 3 hit dice (round up). At level nine, she can cast Bloodwind and Greater Mighty Wallop three times each per day. At level fifteen, she can cast Sharptooth and Razorfangs at will. 
\item Mighty Wings: the Dragonfire Adept gains two Secondary Wing attacks (1d4+Str*0.5 for a Medium character) and a Flight speed of double her land speed (Clumsy). At level nine, she can cast Wings of Cover and Wings of Flurry three times each per day, and the Flight speed becomes (Average), with the option to reduce it to Clumsy but at triple speed. At level fifteen, she can cast Primal Speed and Lord of the Sky at will, and her Flight speed becomes triple her land speed (Good), with the option to reduce it to Clumsy but at x10 speed. 
\item Master of Exhalations: the Dragonfire Adept can cast the following spells at will: Dispelling Breath, Blinding Breath. Additionally, she can leave a number of 5' square "gaps" in her breath area equal to her hit dice, to save allies from death. At level nine she can also cast Rebuking Death and Stunning Breath, three times each per day. At level fifteen, she can also cast Ethereal Breath, Animate Breath, Enervating Breath and Breath Weapon Admixture once each per day. 
\item Disciple of Bahamut: the Dragonfire Adept has a permanent Magic Circle Against Evil effect, can Detect Evil at will, and once per day can make her breath weapon Smiting: the Save DC increases by +4 against Evil targets, and is automatically Empowered. At level nine, she can Smite a second time per day, and can cast Voice of the Dragon and Antidragon Aura three times each per day. At level fifteen, she can Smite a third time per day, and can cast Aspect of the Platinum Dragon twice per day. 
\item Servant of Tiamat: the Dragonfire Adept can cast Mesmerising Glare and Entice Gift at will, and has a permanent Magic Circle Against Good effect. At level nine, she can cast Curse of the Elemental Lords and Trance of the Verdant Domain three times each per day, and once per day can use two different breath weapons at the same time - though they must not share any area. At level fifteen, she can instead use three different breath weapons at the same time, again with no two areas overlapping, and may do so three times per day, and can cast Aspect of the Chromatic Dragon once per day. 
\item Visions of Lichdom: the Dragonfire Adept already knows she wants to be a Dracolich. She has a pseudophylactery: some important item that is used as a focus, from which she can cast Magic Jar at will, but only on reptillian Animals with fewer hit dice than she has. At level nine, she becomes immune to Poison and Disease, and she can choose to release a rotting ichor breath weapon - instead of dealing damage, all in the area must pass a Fortitude Save or be Nauseated for 1 round per 5 dice of damage it deals (round up). At level fifteen, her type changes to Undead, gaining the [Dark Minded] Subtype. She can choose whether or not to also have the [Unliving] Subtype. Protip: Con scores are awesome.
\end{itemize*}

\textbf{Dragon Age (Ex):} at level three, the Dragonfire Adept stops ageing completely. Bonuses and Penalties are no longer gained (and are retroactively lost), she doesn't visibly age, and she won't die of old age. Age-manipulating spells and effects have no effect on her. She is from this point on considered an Adult Dragon, regardless of age. 

\textbf{Explosive Breath (Su):} at level four, the Dragonfire Adept gains the option of releasing her breath as a fireball type effect: it creates a Blast in Medium Range, with a radius of 10' plus 5' per four levels (round up). 

\textbf{Arcane Insight (Su):} at level four, the Dragonfire Adept constantly Detects Magic. All the time. With a Full Round Action she can cast Identify at will. 

\textbf{Shocking Breath (Su):} at level five, the Dragonfire Adept gains the option to deal Electricity damage with her Breath Weapon. 

\textbf{Dragoncraft (Ex):} the fifth-level Dragonfire Adept can simply make stuff out of dead dragons as though it were no big deal. She effectively has Craft Magic Arms and Armour, but only for the purpose of making magic Dragonscale Armour and Shields, or weapons from dragon bones. She needn't supply any actual spells or experience for this, merely the dragon parts, half the item value in components, and time. 

\textbf{Sphere Access:} at level six, the Dragonfire gains Basic Access to an Elemental Sphere. At level twelve, this improves to Advanced Access, and she gains Basic Access to another Elemental or Fiendish Sphere. At level eighteen, she gains Basic Access to another Elemental or Fiendish Sphere, and the first Sphere's access upgrades to Expert. 

\textbf{Radiating Breath (Su):} at level seven, the Dragonfire Adept gains the ability to release her breath from every pore in her body, unleashing it as an Emanation centred on her but not actually including her. It extends out to Close Range. 

\textbf{Scales I (Ex):} at level seven, the Dragonfire Adept develops special scales. She gains an Enhancement Bonus to Natural Armour of +1 per 3 hit dice (round up), and gains Resistance to Fire, Cold, Electricity and Acid equal to her hit dice. 

\textbf{Corrosive Breath (Su):} at level eight, the Dragonfire Adept can choose to release a wave of acid or corrosive gas for her breath weapon. It deals Acid damage, and the save for half is a Fortitude Save. 

\textbf{Booming Breath (Su):} at level ten, the Dragonfire Adept can make her breath weapon deal Sonic damage. This can be unleashed either as a massive boom that requires a Fortitude Save for half, or a high pitched keening that requires a Will Save for half. 

\textbf{Exhalation Enhancement (Su):} at level ten, the Dragonfire Adept can add an annoying rider or bonus to the elemental breath weapons:

\begin{itemize*}
\item Fire: targets that fail their saving throw catch fire, make a Dispel attempt against all [Cold] effects in the area 
\item Cold: targets that fail their saving throw are Slowed for one round, all water in the area is subject to a Freezing Sphere effect 
\item Electricity: targets that fail their save are Entangled for one round 
\item Acid: damage completely ignores Hardness and deals the full amount to objects, targets that fail their save are Sickened for one minute 
\item Sonic: the Booming option causes a Shatter effect on all objects, and those who fail the save are Deafened for one minute. Those who fail to save against the Keening option Panic for 1 round.
\end{itemize*}

\textbf{Spell-Like Abilities II (Sp):} at level eleven, the Dragonfire Adept can cast the following once each per day: Dominate Person, Control Winds, Geas, Wall of Stone, Detect Thoughts, Wall of Ice, Plant Growth, Hallucinatory Terrain. The caster level equals her hit dice, and the Save DC is 10 + half her hit dice + her Charisma Bonus. 

\textbf{Concussive Breath (Su):} at level thirteen, the Dragonfire Adept gains the ability to deal Force damage with her breath weapon. This does not benefit from Exhalation Enhancement, though it's super effective against ghosts. 

\textbf{Scales II (Ex):} at level fourteen, the Dragonfire Adept's scales, which were developed from research and study, are even better. The Enhancement Bonus to Natural Armour is also added as a Resistance Bonus to Saving Throws, and the Energy Resistances are doubled. 

\textbf{Shadow Breath (Su):} at level fourteen, the Dragonfire Adept can choose to exhale shadowy darkness. Instead of dealing damage with a save for half, this deals 1 negative level for every 10 dice of damage it would deal (round down), with a Fortitude Save to negate. The negative levels will go away on their own in an hour, but can still be deadly right now. Yes, Empowering the Breath means more dice which can then be converted into more negative levels. 

\textbf{Spell-Like Abilities III (Sp):} at level sixteen, the Dragonfire Adept can cast the following once each per day: Dominate Monster, Control Weather, Justice of the Wyrm King, Mind of the Labyrinth, Scintillating Scales. The caster level equals her hit dice, and the Save DC is 10 + half her hit dice + her Charisma Bonus. 

\textbf{Gaseous Breath (Su):} at level seventeen, the Dragonfire Adept may release a special gas for the breath weapon. It deals no damage, but all in the area must pass a Will Save or have one of the following occur:

\begin{itemize*}
\item Fall Asleep for one round per five damage dice (round up) 
\item Be Paralysed for one round per five damage dice (round up) 
\item Panic for one round per five damage dice (round up)
\end{itemize*}

\textbf{Final Dragon Age (Ex):} at level eighteen, the Dragonfire Adept is considered a Great Wyrm for the purpose of any obscure crap that is based on age categories. This also means she can take Epic Feats as long as she otherwise meets the requirements. 

\textbf{Draconic Immunities (Ex):} at level nineteen, the Dragonfire Adept is immune to Paralysis, Sleep and one Energy Type of her choice. She may choose to, instead of the Energy Type, be immune to something roughly as common like [Compulsion] effects or [Death] effects or whatever. 

\textbf{Annihilating Breath (Su):} at level twenty, the Dragonfire Adept can choose to have her breath weapon deal "damage". This is Typeless and ignores all Damage Reduction, Energy Resistance/Immunity and even Regeneration. If the target is reduced to zero or fewer hit points, they are utterly disintegrated. 

\textbf{Sorcerous Secrets:} at level twenty, the Dragonfire Adept is so awesome that she can actually form Dragon Pacts: as the dragon in the relationship. Also, she knows how to do whatever is necessary to die in a draconic way: she can become the possessing entity of a graveyard thing, or make a Phylactery and become a Dracolich, or just go and get killed by adventurers or whatever.