%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{General Feats}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%
\featentry{Ability Focus}
%%%

\textbf{Prerequisite:} Special attack that allows a save (Breath Weapon, Stunning Attack, etc).

\textbf{Benefit:} Add +2 to the save DC of your selected special attack.

\textbf{Special:} You can take this feat more than once. The effects don't stack. Instead, select a new special attack each time.

%%%
\featentry{Armor Proficiency (Heavy)}
%%%

\textbf{Prerequisites:} \linkfeat{Armor Proficiency (Light)}, \linkfeat{Armor Proficiency (Medium)}.

\textbf{Benefit:} See Armor Proficiency (Light).

\textbf{Normal:} See Armor Proficiency (Light).

%%%
\featentry{Armor Proficiency (Light)}
%%%

\textbf{Benefit:} When you wear a type of armor with which you are proficient, use the armor's normal Armor Check Penalty and Armor Stealth Penalty with your skills, and to determine your movement rate.

\textbf{Normal:} A character who is wearing armor with which she is not proficient counts the Armor Check Penalty and Armor Stealth Penalty as if it were 4 worse than normal when making skill checks or calculating their maximum movement speed.

%%%
\featentry{Armor Proficiency (Medium)}
%%%

\textbf{Prerequisite:} \linkfeat{Armor Proficiency (Light)}.

\textbf{Benefit:} See Armor Proficiency (Light).

\textbf{Normal:} See Armor Proficiency (Light).

%%%
\featentry{Augment Summoning}
%%%

\textbf{Prerequisite:} \linkfeat{Spell Focus} (conjuration).

\textbf{Benefit:} Each creature you conjure with any \textit{summon} spell gains a +4 enhancement bonus to Strength and Constitution for the duration of the spell that summoned it.

%%%
\featentry{Awesome Blow}{[General] [Fighter]}
%%%

\textbf{Prerequisites:} Str 25, size Large or larger.

\textbf{Benefit:} As a standard action, the creature may choose to subtract 4 from its melee attack roll and deliver an awesome blow. If the creature hits a corporeal opponent smaller than itself with an awesome blow, its opponent must succeed on a Reflex save (DC = damage dealt) or be knocked flying 10 feet in a direction of the attacking creature's choice and fall prone. The attacking creature can only push the opponent in a straight line, and the opponent can't move closer to the attacking creature than the square it started in. If an obstacle prevents the completion of the opponent's move, the opponent and the obstacle each take 1d6 points of damage, and the opponent stops in the space adjacent to the obstacle.

%%%
\featentry{Boneblade Master}
%%%

You have mastered the alchemic processes needed to create boneblades, as well as their use in combat.

\textbf{Prerequisite:} Craft (alchemy, weapons) 4 ranks

\textbf{Benefit:} You are considered to be proficient in the use of any weapon made from the special material Boneblade, and you may craft weapons out of Boneblade. In addition, you are considered to have the \linkfeat{Improved Critical} feat for any boneblade weapons you use in melee combat.

You also gain a +2 to Initiative checks when wielding a boneblade weapon.

%%%
\featentry{Child Necromancer}
%%%

An obsession with death and experimentation with necromancy early in your childhood perverted your body and blossoming magical talent. As a result, your body never aged past childhood, and you are an adult in a child's body, magically powerful but physically weak.

\textbf{Prerequisite:} Caster level 1, must know at least one necromancy spell of each spell level you can cast.

\textbf{Benefit:} All Necromancy spells you cast are at +4 caster level, and you gain the effects of Weapon Finesse for all Necromancy touch attack spells you use (if you desire). You have -4 Strength, and appear to be a child despite your actual age category (this does not prevent penalties or bonuses from advancing in age categories, or stop the aging process). You are one size category smaller than normal for your race (do not further adjust ability modifiers). If you are a spontaneous caster, you may permanently exchange any spell known for any Necromancy spell you possess in written or scroll form. If you are a preparation caster, you may learn any Necromancy spell you possess in written or scroll form from any list, and you may not select Necromancy as a restricted school. These Necromancy spells may be from any list, can be exchanged at any time, and once gained are cast as spells of your spellcasting class. These spells remain as spells known even if you later lose this feat.

\textbf{Special:} This feat can only be taken at 1st level. If circumstances ever cause a character to no longer meet the prerequisites of this feat, they may choose any metamagic feat they qualify for to permanently replace this feat.

%%%
\featentry{Eschew Materials}
%%%

\textbf{Benefit:} You can cast any spell that has a material component costing 1 gp or less without needing that component. (The casting of the spell still provokes attacks of opportunity as normal.) If the spell requires a material component that costs more than 1 gp, you must have the material component at hand to cast the spell, just as normal.

%%%
\featentry{Extra Turning}
%%%

\textbf{Prerequisite:} Ability to turn or rebuke creatures.

\textbf{Benefit:} Each time you take this feat, you can use your ability to turn or rebuke creatures four more times per day than normal.

If you have the ability to turn or rebuke more than one kind of creature each of your turning or rebuking abilities gains four additional uses per day.

\textbf{Normal:} Without this feat, a character can typically turn or rebuke undead (or other creatures) a number of times per day equal to 3 + his or her Charisma modifier.

\textbf{Special:} You can gain Extra Turning multiple times. Its effects stack. Each time you take the feat, you can use each of your turning or rebuking abilities four additional times per day.

%%%
\featentry{Hover}
%%%

\textbf{Prerequisite:} Fly speed.

\textbf{Benefit:} When flying, the creature can halt its forward motion and hover in place as a move action. It can then fly in any direction, including straight down or straight up, at half speed, regardless of its maneuverability.

If a creature begins its turn hovering, it can hover in place for the turn and take a full-round action. A hovering creature cannot make wing attacks, but it can attack with all other limbs and appendages it could use in a full attack. The creature can instead use a breath weapon or cast a spell instead of making physical attacks, if it could normally do so.

If a creature of Large size or larger hovers within 20 feet of the ground in an area with lots of loose debris, the draft from its wings creates a hemispherical cloud with a radius of 60 feet. The winds so generated can snuff torches, small campfires, exposed lanterns, and other small, open flames of non-magical origin. Clear vision within the cloud is limited to 10 feet. Creatures have concealment at 15 to 20 feet (20\% miss chance). At 25 feet or more, creatures have total concealment 
(50\% miss chance, and opponents cannot use sight to locate the creature).

Those caught in the cloud must succeed on a Concentration check (DC 10 + 1/2 creature's HD) to cast a spell.

%%%
\featentry{Improved Critical}{[General] [Fighter]}
%%%

Choose one type of weapon.

\textbf{Prerequisite:} Proficient with weapon, base attack bonus +8.

\textbf{Benefit:} When using the weapon you selected, your threat range is doubled.

\textbf{Special:} You can gain Improved Critical multiple times. The effects do not stack. Each time you take the feat, it applies to a new type of weapon.

This effect doesn't stack with any other effect that expands the threat range of a weapon.

%%%
\featentry{Improved Initiative}{[General] [Fighter]}
%%%

\textbf{Benefit:} You get a +4 bonus on initiative checks.

%%%
\featentry{Natural Spell}
%%%

\textbf{Prerequisites:} Wis 13, wild shape ability.

\textbf{Benefit:} You can complete the verbal and somatic components of spells while in a wild shape. You substitute various noises and gestures for the normal verbal and somatic components of a spell.

You can also use any material components or focuses you possess, even if such items are melded within your current form. This feat does not permit the use of magic items while you are in a form that could not ordinarily use them, and you do not gain the ability to speak while in a wild shape.

%%%
\featentry{Rapid Shot}{[General] [Fighter]}
%%%

\textbf{Prerequisites:} Dex 13, \linkfeat{Point Blank Shot}.

\textbf{Benefit:} You can get one extra attack per round with a ranged weapon. The attack is at your highest base attack bonus, but each attack you make in that round (the extra one and the normal ones) takes a -2 penalty. You must use the full attack action to use this feat.

%%%
\featentry{Spell Focus}
%%%

\textbf{Benefit:} Choose a school of magic. Add +1 to the Difficulty Class for all saving throws against spells from the school of magic you select.

\textbf{Special:} You can gain this feat multiple times. Its effects do not stack. Each time you take the feat, it applies to a new school of magic.

%%%
\featentry{Stand Still}
%%%

You can prevent foes from fleeing or closing.

\textbf{Prerequisite:} Str 13.

\textbf{Benefit:} When a foe's movement out of a square you threaten grants you an attack of opportunity, you can give up that attack and instead attempt to stop your foe in his tracks. Make your attack of opportunity normally. If you hit your foe, he must succeed on a Reflex save against a DC of 10 + your damage roll (the opponent does not actually take damage), or immediately halt as if he had used up his move actions for the round.

Since you use the Stand Still feat in place of your attack of opportunity, you can do so only a number of times per round equal to the number of times per round you could make an attack of opportunity (normally just one).

\textbf{Normal:} Attacks of opportunity cannot halt your foes in their tracks.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%



%%%
\featentry{Ghost Cut Technique}

Study of the ephemeral essence of incorporeal undead has taught you combat techniques
that transcend the limitations of the flesh.

\textbf{Prerequisite:} Whispers of the Otherworld

\textbf{Benefit:} Each day, you can use the spell \textit{wraithstrike} as swift
action spell-like ability a number of time equal to half your character level.

You also gain a +2 to initiative checks, a +4 to Move Silently checks, and Lifesight
as a Special Quality.
