%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Leadership Feats}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Not all games allow Leadership feats, because all the additional Followers and the Cohort can be a lot to keep track of. If a Character has more than one Leadership feat they still only gain a single set of Followers, and they can still only have a single Cohort. However, they can mix their Follower and Cohort selection from among any allowed to them by their feats (and/or class features).

%%%
\featentry{Army of Demons}{[Leadership] [Skill] [Fiend] [Celestial]}
%%%

You have an army of planar crazy crap. This feat scales to your Knowledge: Planes skill.

\textbf{Benefits:} You have a Command Rating equal to your Knowledge: Planes ranks divided by five (round up).

\textbf{4:} You can muster a group of followers. Your leadership score is your ranks in Knowledge: Planes plus your Charisma mod. These followers can and must be outsiders.

\textbf{9:} Your followers swell in number to that of an army.

\textbf{14:} You own a planar stronghold.

\textbf{19:} Your allies gain a +2 morale bonus to all saving throws if they can see you and you are within medium range.

%%%
\featentry{Bureaucrat}{[Leadership] [Skill]}
%%%

You have a functioning guild that makes stuff for you and gives you money. This feat scales to your Appraise skill.

\textbf{Benefits:} You draw an income for working as an administrator, getting 1 GP/week per rank in Appraise.

\textbf{4:} You can muster a group of followers. Your leadership score is your ranks in Appraise plus your Intelligence modifier. These followers all have profession and craft skills.

\textbf{9:} You get your own Stronghold.

\textbf{14:} You get a +2 bonus to profit checks.

\textbf{19:} Your guild goes planar, your number of followers swell to the size of an army and their ranks start filling up with producers and managers from other planes of existence.

%%%
\featentry{Command}{[Leadership] [Combat]}
%%%

You lead tiny men.

\textbf{Benefits:} You have a Command Rating equal to your Base Attack Bonus divided by five (round up).

\textbf{+1:} You can muster a group of followers. Your leadership score is your Base Attack Bonus plus your Charisma Modifier.

\textbf{+6:} You are able to delegate command to a loyal cohort. A cohort is an intelligent and loyal creature with a CR at least 2 less than your character level. Cohorts gain levels when you do.

\textbf{+11:} With a Swift Action you may rally troops, allowing all allies within medium range of yourself to reroll their saves vs. Fear and gain a +2 Morale Bonus to attack and damage rolls for 1 minute. This is a language-dependent ability that may be used an unlimited number of times.

\textbf{+16:} Your allies gain a +2 morale bonus to all saving throws if they can see you and you are within medium range.

%%%
\featentry{Lord of Death}{[Leadership] [Skill] [Necromantic]}
%%%

A whole bunch of skeletons and crap show up to fight under your tattered banner.

\textbf{Benefits:} You have a Command Rating equal to your ranks in Knowledge Religion divided by five (round up).

\textbf{4:} You can muster a group of followers. Your leadership score is your ranks in Knowledge Religion plus your Wisdom modifier. Your followers are all mindless Undead. You don't make them or anything, they just show up.

\textbf{9:} You are able to delegate command to a loyal cohort. Your cohort is an intelligent and loyal Undead creature with a CR 2 less than your character level. Cohorts gain levels when you do.

\textbf{14:} Your followers swell in number to that of an army.

\textbf{19:} Your allies gain energy resistance to Positive Energy equal to your level while they are within line of sight of you.

%%%
\featentry{Monster Rancher}{[Leadership] [Skill]}
%%%

You can breed and train a large number of crazy beasts. This feat scales to your ranks in Handle Animal.

\textbf{Benefits:} You can use Handle Animal as if it were Diplomacy when dealing with Magical Beasts and Dragons. You can do similarly with Aberrations and Plants with an Intelligence Score that is less than 9.

\textbf{4:} You can muster a group of followers. Your leadership score is your ranks in Handle Animal plus any synergy bonuses you gt to that skill. Your followers can, and must be monsters.

\textbf{9:} You have a loyal cohort that is a monster of some kind. A cohort is an intelligent and loyal creature with a CR at least 2 less than your character level. Cohorts gain levels when you do.

\textbf{14:} You know what any monster is unless it is disguised by illusion, and you can look up its stat line in the appropriate monster book when devising your strategies.

\textbf{19:} Once per day, you can reroll a saving throw allowed by a Supernatural Ability.

%%%
\featentry{Tyrant}{[Leadership] [Skill]}
%%%
You push people around and get larger and larger groups trapped in the iron gauntlet of your brutal rule. This feat scales with your ranks in Intimidate

\textbf{Benefits:} You inspire such terror that creatures you intimidate continue to act intimidated after you leave, too afraid to raise their voice in defiance even after you have apparently left them far behind.

\textbf{4:} You can muster a group of followers. Your leadership score is your ranks in Intimidate plus your Strength modifier.

\textbf{9:} Your followers swell in number to that of an army.

\textbf{14:} Your mere presence inspires fear and can break a battle. Enemies with more than 5 hit dice less than you do must make a Will save (DC 10 + half Level + Strength Modifier) of flee in panic. This is a [Fear] effect.

\textbf{19:} Your presence causes despair in even brave opponents. All enemies within 30' of your suffer a -2 Morale penalty to Willpower saves.
