%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\raceentry{Gnome}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{itemize*}
\item +2 Constitution, -2 Strength.
\item Small sized.
\item 20ft movement.
\item Low-Light Vision
\item Weapon Familiarity: Gnomes may treat gnome hooked hammers as martial weapons rather than exotic weapons.
\item +2 racial bonus on saving throws against Illusions.
\item Add +1 to the Difficulty Class for all saving throws against illusion spells cast by gnomes. This adjustment stacks with those from similar effects.
\item +1 racial bonus on attack rolls against kobolds and goblinoids.
\item +4 dodge bonus to Armor Class against monsters of the giant type.
\item +2 racial bonus on \linkskill{Listen} checks.
\item +2 racial bonus on \linkskill{Craft} (alchemy) checks.
\item Automatic Languages: Common and Gnome.
\item Bonus Languages: Draconic, Dwarven, Elven, Giant, Goblin, and Orc. In addition, a gnome can speak with a burrowing mammal (a badger, fox, rabbit, or the like, see below). This ability is innate to gnomes. See the \textit{speak with animals} spell description.
\item Spell-Like Abilities: 1/day-- \textit{speak with animals} (burrowing mammal only, duration 1 minute). A gnome with a Charisma score of at least 10 also has the following spell-like abilities: 1/day -- \textit{dancing lights, ghost sound, prestidigitation}. Caster level 1st; save DC 10 + gnome's Cha modifier + spell level.
\item Favored Class: \linkclass{Bard}. A multiclass gnome's bard class does not count when determining whether he takes an experience point penalty.
\end{itemize*}
